The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-01-29 paper/paper.tex
narrow table
more discussion

    method      &   GPVC    &   Type of bespoke covariances used on each basis 
                                function (see \cite{Almosallam2016b} for 
                                the different options)\\

@@

\begin{table*}[]
\begin{tabular}{|l|llll|llll|}
                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
{\sc Autokeras\_LSTM } & 31.1 & 74.9    & 103.4    & nan    & 44.1             & 67.5             & 122.2             & 98.4\\
{\sc CNN } & 27.1 & 50.8    & 76.7    & 103.3    & 30.5             & 57.7             & 95.4             & 122.4\\
{\sc ComplexSOM } & 38.0 & 52.1    & 94.4    & 101.6    & 34.9             & 45.0             & 91.6             & 100.3\\
{\sc ENSEMBLE1 } & 65.6 & 65.6    & 94.8    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
{\sc funbins } & 36.8 & 82.5    & 122.1    & 141.8    & 42.1             & 100.4             & 142.5             & 167.2\\
{\sc GPzBinning } & 26.2 & 49.7    & 80.8    & 111.7    & 27.8             & 55.3             & 87.2             & 126.9\\
{\sc IBandOnly } & 38.0 & 50.0    & 54.4    & 57.3    & 38.0             & 50.0             & 54.4             & 57.3\\
{\sc JaxCNN } & 59.9 & 101.9    & 105.2    & nan    & 79.7             & 125.2             & 150.0             & nan\\
{\sc JaxResNet } & 73.7 & 111.1    & 131.6    & nan    & 82.2             & 126.0             & nan             & 161.5\\
{\sc LGBM } & 27.0 & 50.8    & 78.8    & 108.2    & 30.3             & 57.4             & 92.9             & 125.2\\
{\sc LSTM } & 26.7 & 51.1    & 81.4    & 103.6    & 30.2             & 57.1             & 95.9             & 126.9\\
{\sc mlpqna } & 39.3 & 64.7    & 93.2    & 121.8    & 42.8             & 72.2             & 109.0             & 133.7\\
{\sc myCombinedClassifiers } & 35.3 & 60.5    & 83.5    & 120.4    & 39.4             & 63.9             & 93.1             & 148.9\\
{\sc NeuralNetwork1 } & 76.3 & 117.6    & 135.8    & 0.0    & 81.9             & 132.9             & 158.0             & 0.0\\
{\sc NeuralNetwork2 } & 30.5 & 48.0    & 57.8    & 136.1    & 49.4             & 102.7             & 122.2             & 140.0\\
{\sc PCACluster } & 29.5 & 68.1    & 75.8    & 0.0    & 50.1             & 72.7             & 90.8             & 0.0\\
{\sc Random } & 1.2 & 1.2    & 1.2    & 1.2    & 1.2             & 1.2             & 1.2             & 1.2\\
{\sc RandomForest } & 39.5 & 65.2    & 93.2    & 118.5    & 43.0             & 72.8             & 110.1             & 136.5\\
{\sc SimpleSOM } & 39.5 & 59.9    & 77.9    & 98.5    & 42.2             & 73.2             & 108.0             & 134.8\\
{\sc TensorFlow\_FFNN } & 39.0 & 64.3    & 90.3    & 114.7    & 43.0             & 72.0             & 112.3             & 137.2\\
{\sc TCN } & 40.0 & 65.3    & 90.9    & 110.5    & 44.1             & 72.8             & 108.0             & 134.0\\
{\sc UTOPIA } & 39.0 & 63.7    & 90.8    & 112.2    & 43.1             & 74.0             & 112.8             & 139.3\\
{\sc ZotBin } & 64.8 & 106.4    & 121.8    & 135.3    & 77.6             & 120.0             & 141.8             & 154.7\\
{\sc ZotNet } & 73.6 & 111.2    & 131.8    & 145.9    & 83.7             & 128.5             & 150.1             & 167.2\\
\end{tabular}
\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the 3x2pt DETF figure-of-merit achieved by entrant methods on the 
CosmoDC2 part of the challenge.}
\end{table*}

\begin{itemize}
    \item Big table of results
    \item fisher matrix ellipses overplotted for top N methods
    \item Plots of n(z)
\end{itemize}

@@

\subsection{Individual winners}

RUN WINNERS WITH UGRIZY TO GET AN IDEA OF THE LOSS FROM DROPPING U AND Y.

\subsection{Large bin numbers}

It is clear from values above that fears about restricted photometry catastrophically limiting
the information needed for tomography are unfounded.  In both the CosmoDC2 and Buzzard analyses,
even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
tomographic bins.
Once bins become this narrow the primary analysis concern will become the calibration of overall
bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.

@@

>>>>

    method      &   GPVC    &   Type of bespoke covariances used on each basis 
                                function (see \citealp{Almosallam2016b})\\

@@

\begin{table*}[]
\begin{tabular}{|l|llll|llll|}
                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
{\sc Autokeras\_LSTM } & 31.1 & 74.9    & 103.4    & nan    & 44.1             & 67.5             & 122.2             & 98.4\\
{\sc CNN } & 27.1 & 50.8    & 76.7    & 103.3    & 30.5             & 57.7             & 95.4             & 122.4\\
{\sc ComplexSOM } & 38.0 & 52.1    & 94.4    & 101.6    & 34.9             & 45.0             & 91.6             & 100.3\\
{\sc ENSEMBLE1 } & 65.6 & 65.6    & 94.8    & 0.0    & 0.0             & 0.0             & 0.0             & 0.0\\
{\sc funbins } & 36.8 & 82.5    & 122.1    & 141.8    & 42.1             & 100.4             & 142.5             & 167.2\\
{\sc GPzBinning } & 26.2 & 49.7    & 80.8    & 111.7    & 27.8             & 55.3             & 87.2             & 126.9\\
{\sc IBandOnly } & 38.0 & 50.0    & 54.4    & 57.3    & 38.0             & 50.0             & 54.4             & 57.3\\
{\sc JaxCNN } & 59.9 & 101.9    & 105.2    & nan    & 79.7             & 125.2             & 150.0             & nan\\
{\sc JaxResNet } & 73.7 & 111.1    & 131.6    & nan    & 82.2             & 126.0             & nan             & 161.5\\
{\sc LGBM } & 27.0 & 50.8    & 78.8    & 108.2    & 30.3             & 57.4             & 92.9             & 125.2\\
{\sc LSTM } & 26.7 & 51.1    & 81.4    & 103.6    & 30.2             & 57.1             & 95.9             & 126.9\\
{\sc mlpqna } & 39.3 & 64.7    & 93.2    & 121.8    & 42.8             & 72.2             & 109.0             & 133.7\\
{\sc myCombinedClassifiers } & 35.3 & 60.5    & 83.5    & 120.4    & 39.4             & 63.9             & 93.1             & 148.9\\
{\sc NeuralNetwork1 } & 76.3 & 117.6    & 135.8    & 0.0    & 81.9             & 132.9             & 158.0             & 0.0\\
{\sc NeuralNetwork2 } & 30.5 & 48.0    & 57.8    & 136.1    & 49.4             & 102.7             & 122.2             & 140.0\\
{\sc PCACluster } & 29.5 & 68.1    & 75.8    & 0.0    & 50.1             & 72.7             & 90.8             & 0.0\\
{\sc Random } & 1.2 & 1.2    & 1.2    & 1.2    & 1.2             & 1.2             & 1.2             & 1.2\\
{\sc RandomForest } & 39.5 & 65.2    & 93.2    & 118.5    & 43.0             & 72.8             & 110.1             & 136.5\\
{\sc SimpleSOM } & 39.5 & 59.9    & 77.9    & 98.5    & 42.2             & 73.2             & 108.0             & 134.8\\
{\sc TensorFlow\_FFNN } & 39.0 & 64.3    & 90.3    & 114.7    & 43.0             & 72.0             & 112.3             & 137.2\\
{\sc TCN } & 40.0 & 65.3    & 90.9    & 110.5    & 44.1             & 72.8             & 108.0             & 134.0\\
{\sc UTOPIA } & 39.0 & 63.7    & 90.8    & 112.2    & 43.1             & 74.0             & 112.8             & 139.3\\
{\sc ZotBin } & 64.8 & 106.4    & 121.8    & 135.3    & 77.6             & 120.0             & 141.8             & 154.7\\
{\sc ZotNet } & 73.6 & 111.2    & 131.8    & 145.9    & 83.7             & 128.5             & 150.1             & 167.2\\
\end{tabular}
\caption{INCOMPLETE, AND AUTO-GENERATED SO NO POINT EDITING! Values of the 3x2pt DETF figure-of-merit achieved by entrant methods on the 
CosmoDC2 part of the challenge.}
\label{tab:cosmodc2}
\end{table*}

\begin{itemize}
    \item fisher matrix ellipses overplotted for top N methods
    \item Plots of n(z)
\end{itemize}

@@

-

\subsection{General outcomes}
It is clear from values above that fears about restricted photometry catastrophically limiting
the information needed for tomography are unfounded.  In both the CosmoDC2 and Buzzard analyses,
even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration of overall
bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.

@@

Methods using fixed bins with equal numbers by nominal redshift tend to plateau at scores of
120-140 in Table \ref{tab:cosmodc2}, suggesting that a range of machine learning models can
effectively classify by bin, at least in this case of representative, extensive training
information.  Notably, this is the case even for the UTOPIA entry, which uses the nearest 
training set neighbour to assign bins and thus was explicitly designed to work only
for this idealised scenario; it should give an upper limit when using the same fixed bins.
That other methods with the same nominal bin choice acheive scores close to it once again
confirms that many entries are close to the best possible score for this approach.

Several methods that instead try to optimize nominal bin edges break this
barrier and acheive very high scores.  This included ZotNet, JaxResNet, and JaxCNN.  The
Stacked Generalization method similarly included manually tuned bin edges to maximize
score for this scenario. The typical 15--20\% improvement in FoM scores when optimizing
nominal bin edges makes clear that this easy win should is compelling.

Notably, though, the method that received the highest overall score in Table \ref{tab:cosmodc2}, 
{\sc funbins} did so with a simpler approach - it used a random forest to assign galaxies to bins,
but chose nominal bin edges using an equal width bins in \emph{comoving distance} $\chi$, converting
this to redshift using a fiducial cosmology. The success of this simple solution suggests this
as an easy heuristic for well-optimized nominal edges.
